

# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r echo=FALSE, message=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r echo=FALSE}
#List of selected countries
selected_countries<-c("Malawi","Ghana", "Cameroon")
#
filtered_df<-data_cmnn %>% 
  filter(Entity %in% selected_countries) %>% rename(daly = 4)
# Pivotin the data to show each country in a separate column.
pivot_df<- pivot_wider(filtered_df,names_from = Entity,values_from = daly,id_cols = Year)
 # Use kable() from the knitr package to render the table.              
   kable(pivot_df) 
```

## Plot Showing Trends in CMNN Burden Over Time

```{r echo=FALSE}
cmnn_plot <- 
  ggplot() +
  geom_line(data = pivot_df, aes(x = Year, y = Cameroon, color = "Cameroon")) +
  geom_line(data = pivot_df, aes(x = Year, y = Ghana, color = "Ghana")) +
  geom_line(data = pivot_df, aes(x = Year, y = Malawi, color = "Malawi")) +
  labs(title = "DALY Burden Over Time",
       x = "Year",
       y = "DALY",
       color = "Country") +
  scale_color_manual(values = c("Cameroon" = "blue", "Ghana" = "red", "Malawi" = "green")) +
  theme_minimal()

cmnn_plot
```

## Summary of CMNN Burden Findings

Malawi had the highest burden among the three countries considered in this analysis from 1990 to 2012. After 2012, Cameroon became the country with the highest burden of the three. Ghana has constantly maintained the lowest burden among the three countries thoughout the period of analysis.




