# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse,knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries

#List of selected countries
selected_countries<-c("Malawi","Ghana", "Cameroon")
#
filtered_df<-data_ncd %>% 
  filter(Entity %in% selected_countries) %>% rename(daly = 4)
# You should pivot the data to show each country in a separate column.
pivot_df<- pivot_wider(filtered_df,names_from = Entity,values_from = daly,id_cols = Year)
 # Use kable() from the knitr package to render the table.              
   kable(pivot_df)                                                             
```

## Plot Showing Trends in NCD Burden Over Time 

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# Load necessary library
library(ggplot2)

ggplot(filtered_df, aes(x = Year, y = daly, color = Entity)) +
  geom_line() +
  geom_point() + 
  labs(title = "Trend of DALY Burden Over Time",
       x = "Year",
       y = "DALY Burden",
       color = "Entity") +  # Update legend title
  scale_color_manual(values = c("Malawi" = "blue", "Ghana" = "red", "Cameroon" = "green")) +  # Set custom colors for each country
  theme(legend.position = "bottom")+  # Change legend position
geom_line()
  
```


## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
The DALY burden of non- communicable diseases rate in Malawi was high between 1990 and 2000 compared to the other two countries. However, there was a steep and continuous decline in non-communicable disease rates from 2000 to 2010. The plot also show a more linear pattern of change in Ghana with lower DALY burden compared to the other countries.



